SHIV: Reducing Supervisor Burden using Support Vectors to Efficiently Learn Robot Control Policies from Demonstrations [v21]

### 2015-09-14 paper.tex
Anca's comments

\newboolean{include-notes}
\setboolean{include-notes}{false}
\newcommand{\adnote}[1]{\ifthenelse{ \boolean{include-notes}}%
 {\textcolor{blue}{\textbf{AD: #1}}}{}}

@@

\begin{abstract}
Online learning from demonstration algorithms, such as DAgger, can learn policies for problems where the system
dynamics and the cost function are unknown. However, during learning they impose a burden on supervisors to provide labels (control
signals) as the robot encounters new states when executing its current best policy. We introduce the SHIV
algorithm (Svm-based reduction in Human InterVention), which reduces this burden by only requesting supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to non-stationary high dimensional state distributions.  To facilitate scaling and outlier rejection, risk is based on distance to an approximate level set boundary defined by a One Class support vector machine.  We compare SHIV with DAgger in three policy learning experiments: 1) a driving simulator with a 27936 dimension visual HOG feature state representation, 2) 4DOF robot arm push-grasping in clutter simulation with Box2D with a 22 dimension state representation  of joint angles and the pose of each object, and 3) surgical needle insertion with a 16 dimension state represented spatial transform.  Results suggest that SHIV can learn equivalent policies with up to $70\%$ fewer queries.

@@

In model-free robot learning from demonstration, a robot learns to perform a task, such as driving or grasping an object in a cluttered environment, from examples provided by a  supervisor, usually a human.  In such problems, the robot does not have access to either the cost function that it should optimize, nor the dynamics model. The former occurs when it is difficult to specify how to trade-off various aspects that matter, like when the robot is reaching for a goal object while not pushing objects off a table.  The latter occurs when either the system or the interaction with the world is difficult to characterize based on observations, as in Fig. \ref{fig:teaser}(a) the robot drives on the track but only has access to visual information from a bird's eye view.

Rather than explicitly learning the cost function, like in inverse reinforcement learning, and the dynamics model, and then using optimization to produce a policy for the task, the robot learns the policy directly from supervisor examples mapping states to controls~\cite{argall2009survey}. Learning from demonstrations algorithms can be categorized as offline, where the robot only observes demonstrations, or online, where the robot interacts with the world and receives feedback from the supervisor.

@@

\includegraphics[width=0.5\textwidth]{figures/teaser.pdf}
\caption{ 
SHIV  reduces the number of queries to the supervisor in our three test domains: a) a driving simulator where the goal is to learn a controller on HOG features extracted from synthetic images to keep a car on a polygonal track; b) push-grasping in clutter in a physics simulator, where the goal is to learn a controller from human demonstrations to grasp the green object while not letting the red square objects fall off the gray boundary representing a table; c) learning a controller for the Da Vinci Research Kit from a master surgeon's demonstrations of correcting a needle before suturing in surgery; d) cost of SHIV and DAgger's policy on the grasping in clutter example versus states labeled.
}
\vspace*{-10pt}
\label{fig:teaser}
\end{figure}

In offline learning from demonstration, the robot learns the policy based on a batch of examples, and then executes it to achieve the task.  During execution, a small error can accumulate, leading the robot away from the region of the state space where it was given examples, leading to unrecoverable failures.  For example, a robot driving may be trained on examples driving safely down the center of a lane, but even slight deviations will put the robot into states near the side of the road where its policy could  fail~\cite{pomerleau1989alvinn}. Ross and Bagnell showed the number of errors made by the robot, in the worst case,  can scale quadratically with the time horizon of the task~\cite{ross2010efficient}.

Online learning from demonstration addresses this issue by iteratively gathering more examples from the supervisor in states the robot encounters~\cite{grollman2007dogged,ross2010efficient,ross2010reduction}. One such algorithm, DAgger learns a series of policies. At each iteration, the robot trains a policy based on the existing examples, then rolls out (executes) that policy and the supervisor provides demonstrations for all states the robot visits. The new state/control examples are aggregated with the old examples for the next iteration. DAgger and related algorithms have been applied in a wide range of applications, from quadrotor flight to natural language to Atari games~\cite{NIPS2014_5421,duvallet2013imitation,ross2013learning}. In DAgger, it was proved under certain conditions, the number of errors scales only linearly with the time horizon of the task.

DAgger significantly increases the burden on the supervisor, who now labels all states that the robot visits during training.  Our goal is to reduce supervisor burden without degrading robot performance.

@@

\section{Related Work}
Below we summarize related work in active approaches to robot learning from demonstration and risk evaluation techniques..

\noindent\textbf{Active Learning from Demonstration}
Online learning from demonstration can be formulated as a stream-based active learning problem ~\cite{atlas1990training,cohn1994improving}. In stream based active learning, the decision of whether to query the supervisor (for a label or control signal) or not is not over the entire state space (like in traditional pool-based active learning), but on states drawn one at a time from some data source. In  online learning  from demonstration, the data source is the states encountered during learning.

Stream-based active learning has two ingredients: a stream, given here by executing the current best policy, and a query selection method deciding whether or not to query. Typical query selection methods are estimator-centric:  evaluating risk using the estimator, e.g. distance from the classification hyperplane \cite{tong2002support}, as in Fig. \ref{fig:support_example}(a). An alternative to distance-based measures of  risk is query by committee  \cite{breiman1996bagging}, which uses a committee of hypothesized estimators that are trained on different subsets of the training data. Risk here is based on the level of agreement or disagreement among these hypotheses, with higher levels of disagreement for a state leading to higher chances of querying that state (Fig. \ref{fig:support_example}(b)).

Both approaches implicitly assume a stationary state distribution --- that the new data is sampled from the same distribution as the previous training data. Although such methods have been previously proposed for online learning from demonstration (see \cite{chernova2009interactive,grollman2007dogged} for the former and \cite{judah2011active,judah2012active} for the latter),
online learning from demonstration violates the stationary distribution assumption because each new policy induces a new distribution of states. This can have negative consequences for learning: it has been shown that when training and test distributions are different, query by committee can perform worse than randomly selecting  which states to query \cite{burbidge2007active}.

The  estimator is a function of the current distribution. Therefore, it no longer provides a good measure of confidence when that distribution changes, as it does when the robot is rolling out a learned policy and starts encountering further away states. Instead of relying on the estimator, our measure of risk explicitly identifies when states are drawn from a different distribution, i.e. when they are novel.

@@

\noindent\textbf{Risk via Novelty Detection.}
One part of our risk definition is the notion that a trained model will be able to generalize within the distribution it is
trained on \cite{tokdar2010importance}, making states outside of this distribution risky. Novelty detection \cite{hodge2004survey} is concerned with recognizing when this happens: recognizing that a sample is outside of this distribution.

@@

@@

\noindent\textbf{Modeling Choices and Assumptions}  We model the system dynamics as Markovian, stochastic, and stationary. Stationary dynamics meaning  given a state and a control the probability of the next state doesn't change over time. Note this is different then the non-stationary distribution over the states the robot encounters during learning. We model the initial state as sampled from a distribution over the state space.

We assume a known state space and set of controls. We assume access to a robot or simulator, such that we  can sample from the state sequences induced by a sequence of controls.   Lastly, we assume access to a supervisor who can, given a state, provide a control signal label. We assume the supervisor can be noisy and imperfect. Noting that the robot cannot surpass the level of performance of the supervisor.

\noindent\textbf{Policies and State Densities.}
Following conventions from control theory, we denote by $\mathcal{X}$ the set consisting of observable states for a robot task, consisting, for example, of \mlnote{We should discuss if we are going to consider images states or not}
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of allowable control inputs for the robot, which can be discrete or
continuous. We model dynamics, as Markovian, such that the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$: 
$$p(\bx_{t+1}|\bu_{t},\bx_{t}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{t+1}|\bu_{t}, \bx_t)$$
We assume a probability density over initial states $p(\bx_0)$.
%We denote the probability density over the initial state also by $p:\mathcal{X}\to \mathbb{R}$.

@@

\noindent\textbf{Objective.} The objective of  policy learning is to find a policy that minimizes some known cost function $C(\hat{\tau}) = \sum^T_{t=1} c(\bx_t,\bu_t)$ of a trajectory $\hat{\tau}$. The cost function $c:\mathcal{X}\times \mathcal{U}\to \mathbb{R}$ is typically user defined and task specific. 
For example, in the task of inserting a peg into a hole, a function on the distance between the peg's current and desired final state can
be considered \cite{levine2015end}.

However in this case, we do not have access to the cost function itself. Instead, we only have access to 
a supervisor that can achieve a desire level of performance on the task.  We are furthermore given
an initial set of $N$ stochastic demonstration trajectories $\lbrace \tilde{\tau}^1,...,\tilde{\tau}^N \rbrace$. 
which are the result of the supervisor applying this policy. This induces a training dataset $\mathcal{D}$ of all state-control input pairs from the demonstrated trajectories.

We define a `surrogate' loss function $l:\mathcal{U}\times \mathcal{U}\to \mathbb{R}$, which provides a distance
measure between any pair of control values. In the continuous case, we consider $l(\bu_0,\bu_1) = ||\bu_0-\bu_1||^2$,
while in the discrete case $l(\bu_0,\bu_1) = 1$ if $\bu_0 \neq \bu_1$ and $l(\bu_0, \bu_1)=0$ otherwise.

@@

 \vspace{-2ex}
\begin{align}\label{eq:LFD_obj}
\underset{\theta}{\min} \: E_{p(\bx|\theta)} [l(\pi_\theta(\bx),\tilde{\pi}(\bx))]
\end{align}
with $p(\bx|\theta)$ from Eq. \ref{eq:density}.

@@

\section{Risky and Safe States}

\begin{figure*}[t]
\centering

@@

Providing correct control inputs or ``labeling'' for all states encountered at each
iteration can impose a large burden on the supervisor.
Instead of asking the supervisor for labels at all visited states, SHIV uses a measure of risk to actively decide whether a label is necessary.

@@

We then  compare our query selection method with those typically used in active learning. We show that for a non-stationary state distribution like ours, the notion of risk based on novelty and misclassified regions performs better than a confidence and query-by-committee based methods. We continue with a sensitivity analysis, which suggests that the performance of SHIV is robust to the choice of how risky the robot is allowed to be (the $\sigma$ parameter from Eq. \ref{eq:decision_func}).

@@

\subsubsection{Comparison to active learning approaches.}\

>>>>

\newboolean{include-notes}
\setboolean{include-notes}{true}
\newcommand{\adnote}[1]{\ifthenelse{ \boolean{include-notes}}%
 {\textcolor{blue}{\textbf{AD: #1}}}{}}

@@

\begin{abstract}
Online learning from demonstration algorithms, such as DAgger, can learn policies for problems where the system
dynamics and the cost function are unknown. However, during learning, they impose a burden on supervisors to provide labels (control
signals) as the robot encounters new states while executing its current best policy. We introduce the SHIV
algorithm (Svm-based reduction in Human InterVention), which reduces this burden by only requesting supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to non-stationary high dimensional state distributions.  To facilitate scaling and outlier rejection, risk is based on distance to an approximate level set boundary defined by a One Class support vector machine.  We compare SHIV with DAgger in three policy learning experiments: 1) a driving simulator with a 27936 dimension visual HOG feature state representation, 2) a push-grasping in clutter simulation for a 4DOF robot arm, using Box2D, with a 22 dimension state representation  of joint angles and the pose of each object, and 3) a surgical needle insertion task with a 16 dimension state represented spatial transform.  Results suggest that SHIV can learn equivalent policies with up to 70\% fewer queries.

@@

In model-free robot Learning from Demonstration (LfD), a robot learns to perform a task, such as driving or grasping an object in a cluttered environment, from examples provided by a  supervisor, usually a human.  In such problems, the robot does not have access to either the cost function that it should optimize, nor the dynamics model. The former occurs when it is difficult to specify how to trade-off various aspects that matter, like when the robot is reaching for a goal object while not pushing objects off a table.  The latter occurs when either the system or the interaction with the world is difficult to characterize based on observations, as in Fig. \ref{fig:teaser}(a) the robot drives on the track but only has access to visual information from a bird's eye view.\adnote{these are both pretty bad examples; i'd use the typical driving example for the first point, taken from abeeel and ng, and i'd use the pushing for the second point, because it is very hard to write down accurate dynamics models for pushing -- took Mehmet 3 years to do push grasping with a single object, still can't do with arbitrary objects; Jen uses Box2d but the real world is uncertain; it's a perfect example in whihc writing down the dynamics model is very difficult;}

Rather than explicitly learning the cost function, like in Inverse Reinforcement Learning\adnote{cite}, and the dynamics model, and then using optimization to produce a policy for the task, in model-free LfD the robot learns the policy directly from supervisor examples mapping states to controls~\cite{argall2009survey}.

@@

\includegraphics[width=0.5\textwidth]{figures/teaser.pdf}
\caption{ 
SHIV  reduces the number of queries to the supervisor in our three test domains: a) a driving simulator where the goal is to learn a controller on HOG features extracted from synthetic images to keep a car on a polygonal track; b) push-grasping in clutter in a physics simulator, where the goal is to learn a controller from human demonstrations to grasp the green object while not letting the red square objects fall off the gray boundary representing a table; c) learning a controller for the Da Vinci Research Kit from a master surgeon's demonstrations of correcting a needle before suturing in surgery; d) normalized performance of SHIV and DAgger's policy on the grasping in clutter example versus states labeled \adbote{by the supervisor}.
}
\vspace*{-10pt}
\label{fig:teaser}
\end{figure}

LfD algorithms can be categorized as offline, where the robot only observes demonstrations, or online, where the robot interacts with the world and receives feedback from the supervisor.
In offline LfD, the robot learns the policy based on a batch of examples, and then executes it to achieve the task.  During execution, a small error can accumulate, leading the robot away from the region of the state space where it was given examples, leading to unrecoverable failures.  For example, a robot driving may be trained on examples driving safely down the center of a lane, but even slight deviations will \adnote{eventually} put the robot into states near the side of the road where its policy could  fail~\cite{pomerleau1989alvinn}. Ross and Bagnell showed the number of errors made by the robot, in the worst case,  can scale quadratically with the time horizon of the task~\cite{ross2010efficient}.

Online LfD addresses this issue by iteratively gathering more examples from the supervisor in states the robot encounters~\cite{grollman2007dogged,ross2010efficient,ross2010reduction}. One such algorithm, DAgger, learns a series of policies. At each iteration, the robot trains a policy based on the existing examples, then rolls out (executes) that policy, and the supervisor provides demonstrations for all states the robot visits. The new state/control examples are aggregated with the old examples for the next iteration. DAgger and related algorithms have been applied in a wide range of applications, from quadrotor flight to natural language to Atari games~\cite{NIPS2014_5421,duvallet2013imitation,ross2013learning}. In DAgger, \adnote{delted it was proved, unnecessary} under certain conditions, the number of errors scales only linearly with the time horizon of the task\adnote{cite}.

\adnote{there needs to be a lead up to this, like yay dagger is cool. however, }DAgger significantly increases the burden on the supervisor, who now labels all states that the robot visits during training.  Our goal is to reduce supervisor burden without degrading robot performance.

@@

\adnote{nothing about experiements and resutls? how do i know if it worked? also how do i know that you didn;t just compare to dagger, but ran an indepth analysis of the algorithm? i thought i put a paragraph like that here at some point}

\section{Related Work}
Below we summarize related work in active approaches to robot learning from demonstration and risk evaluation techniques.

\noindent\textbf{Active Learning from Demonstration}
Online LfD\adnote{been replacing with "LfD" for simplicity, no need to say the same expression 100 times in the paper especially since everyone calls it that} can be formulated as a stream-based active learning problem ~\cite{atlas1990training,cohn1994improving}\adnote{NO!! Your approach of asking only when risky instantiates stream based active learning. In general, online lfd is NOT an active learning problem, that is the proble we are fixing here! sorry for the "!!" but this is crucial}. In stream based active learning, the decision of whether to query the supervisor (for a label or control signal) or not is not over the entire state space (like in traditional pool-based active learning), but on states drawn one at a time from some data stream. In  online LfD, the\adnote{this} data stream is the states encountered during learning\adnote{encountered when the robot is executing the current best policy.}

Stream-based active learning has two ingredients: a stream, given here by executing the current best policy,\adnote{Aside from the data strea, stream-based active learning has another ingredient: a query selection method.} and a query selection method deciding whether or not to query. Typical query selection methods are estimator-centric, evaluating risk using the estimator, e.g. distance from the classification hyperplane \cite{tong2002support}, as in Fig. \ref{fig:support_example}(a); or query by committee  \cite{breiman1996bagging}, which uses a committee of hypothesized estimators that are trained on different subsets of the training data. Risk in query by committee is based on the level of agreement or disagreement among these hypotheses, with higher levels of disagreement for a state leading to higher chances of querying that state (Fig. \ref{fig:support_example}(b)).

Both approaches implicitly assume a stationary state distribution --- that the new data is sampled from the same distribution as the previous training data. Although such methods have been previously proposed for online LfD (see \cite{chernova2009interactive,grollman2007dogged} for the former and \cite{judah2011active,judah2012active} for the latter),
online LfD violates the stationary distribution assumption because each new policy induces a new distribution of states. This can have negative consequences for learning: it has been shown that when training and test distributions are different, query by committee can perform worse than randomly selecting  which states to query \cite{burbidge2007active}.

The problem lies in being estimator-centric. The  estimator is a function of the current distribution. Therefore, it no longer provides a good measure of confidence when that distribution changes, as it does when the robot is rolling out a learned policy and starts encountering further away states. Instead of relying on the estimator, our measure of risk explicitly identifies when states are drawn from a different distribution, i.e. when they are novel.

@@

\adnote{replace learning from demonstration with LfD from here onwards}

\noindent\textbf{Risk via Novelty Detection.}
The main motivation for our risk definition is the notion that a trained model will be able to generalize within the distribution it is
trained on \cite{tokdar2010importance}, making states outside of this distribution risky. Novelty detection \cite{hodge2004survey} is concerned with recognizing when this happens: recognizing that a sample is outside of this distribution.

@@

\adnote{OK fiiiiiine. Then maybe To our knowledge, incorporating a query selection method for nonstationary distributions, such as the One Class SVM, into online LfD is a novel contribution.}

@@

\noindent\textbf{Modeling Choices and Assumptions}  We model the system dynamics as Markovian, stochastic, and stationary. Stationary dynamics meaning\adnote{this sentence is missing a predicate; Stationary dynamics occur when, given..}  given a state and a control, the probability of the next state doesn't\adnote{does not} change over time. Note this is different then\adnote{from} the non-stationary distribution over the states the robot encounters during learning. \adnote{very good addition here!}

We model the initial state as sampled from a distribution over the state space.
We assume a known state space and set of controls. We also assume access to a robot or simulator, such that we  can sample from the state sequences induced by a sequence of controls.   Lastly, we assume access to a supervisor who can, given a state, provide a control signal label. We assume the supervisor can be noisy and imperfect, noting that the robot cannot surpass the level of performance of the supervisor. \adnote{not in general just in our work}

\adnote{again, this would be better at the same time with introducing the basic notation, like states and controls, so that you can state mathematically what these mean; for instance, you restate the markovian modelling anyway}

\noindent\textbf{Policies and State Densities.}
Following conventions from control theory, we denote by $\mathcal{X}$ the set consisting of observable states for a robot task, consisting, for example, of \mlnote{We should discuss if we are going to consider images states or not}
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of allowable control inputs for the robot, which can be discrete or
continuous. We model dynamics as Markovian, such that the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$: 
$$p(\bx_{t+1}|\bu_{t},\bx_{t}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{t+1}|\bu_{t}, \bx_t)$$
We assume a probability density over initial states $p(\bx_0)$.
%We denote the probability density over the initial state also by $p:\mathcal{X}\to \mathbb{R}$.

@@

\noindent\textbf{Objective.} The objective of  policy learning is to find a policy that minimizes some known cost function $C(\hat{\tau}) = \sum^T_{t=1} c(\bx_t,\bu_t)$ of a trajectory $\hat{\tau}$. The cost $c:\mathcal{X}\times \mathcal{U}\to \mathbb{R}$ is typically user defined and task specific. 
For example, in the task of inserting a peg into a hole, a function on the distance between the peg's current and desired final state can
be considered \cite{levine2015end}.  \adnote{can be considered? like you'll allow it? maybe For example, in the task... , the cost function could be ...}

In our problem, we do not have access to the cost function itself. Instead, we only have access to 
a supervisor that can achieve a desire\adnote{d} level of performance on the task. \adnote{furthermore doesnt make sense here because the supervisor is providing these; say instead The supervisor provides the robot an initial set..} We are furthermore given
an initial set of $N$ stochastic demonstration trajectories $\lbrace \tilde{\tau}^1,...,\tilde{\tau}^N \rbrace$. 
which are the result of the supervisor applying this policy. This induces a training data set $\mathcal{D}$ of all state-control input pairs from the demonstrated trajectories.

We define a `surrogate' loss function \adnote{as in \cite{ross2010reduction}, }$l:\mathcal{U}\times \mathcal{U}\to \mathbb{R}$, which provides a distance
measure between any pair of control values. In the continuous case, we consider $l(\bu_0,\bu_1) = ||\bu_0-\bu_1||^2$,
while in the discrete case $l(\bu_0,\bu_1) = 1$ if $\bu_0 \neq \bu_1$ and $l(\bu_0, \bu_1)=0$ otherwise.

@@

 \vspace{-2ex}
\begin{align}\label{eq:LFD_obj}
\underset{\theta}{\min} \: E_{p(\bx|\theta)} [l(\pi_\theta(\bx),\tilde{\pi}(\bx))]
\end{align}
with $p(\bx|\theta)$ from Eq. \ref{eq:density}.\adnote{fix}

@@

\section{Risky and Safe States}
\adnote{I don't like the flow of this coming before the algorithm. You can understand the algorithm without understanding the specifics of risk, but here I read this and still don't know how the algorithm is going to work and it's like keeping me in (bad) suspense until the next section. It also breaks flow because the last section set up the optimization problem that shiv and dagger are solving}
\begin{figure*}[t]
\centering

@@

Providing correct control inputs for (or ``labeling'')  all states encountered at each
iteration can impose a large burden on the supervisor.\adnote{but you haven't described the algorithm yet so how am i supposed to know what that means; this is no longer intro land so stuff should be precise not handwavy; that is why this section doesnt fit here}
Instead of asking the supervisor for labels at all visited states, SHIV uses a measure of risk to actively decide whether a label is necessary.

@@

We then  compare our query selection method with those typically used in active learning. We show that for a non-stationary state distribution like ours, the notion of risk based on novelty and misclassified regions performs better than a confidence and query-by-committee based methods.\adnote{also: We also test whether SHIV actually benefits from measuring both forms of risk (novelty and ???).} We continue with a sensitivity analysis, which suggests that the performance of SHIV is robust to the choice of how risky the robot is allowed to be (the $\sigma$ parameter from Eq. \ref{eq:decision_func}).

@@

\subsubsection{Comparison to active learning approaches.}
